---
title: "HW06"
author: "Zixin Ouyang"
date: "10/26/2017"
output:
  pdf_document: default
  html_document: default
---

```{r}
wisc_train<-read.csv('wisc-trn.csv')
wisc_test<-read.csv('wisc-tst.csv')
```

```{r}
library(caret)
```

## Exercise 1
```{r}
set.seed(1337)
knn_mod = train(
  class ~ .,
  data = wisc_train,
  method = "knn",
  trControl = trainControl(method = "cv", number = 5),
  tuneGrid = expand.grid(k = seq(1, 101, by = 2))
)
```

```{r}
plot(knn_mod)
```

## Exercise 2
```{r}
set.seed(1337)
knn_mod2 = train(
  class ~ .,
  data = wisc_train,
  method = "knn",
  trControl = trainControl(method = "cv", number = 5),
  preProcess = c("center", "scale"),
  tuneGrid = expand.grid(k = seq(1, 101, by = 2))
)
```

```{r}
plot(knn_mod2)
```

## Exercise 3
```{r, message=FALSE, warning=FALSE}
library(randomForest)
```

```{r}
set.seed(1337)
rf_mod = train(
  class ~ .,
  data = wisc_train,
  method = "rf",
  trControl = trainControl(method = "cv", number = 5),
  tuneGrid = expand.grid(mtry = seq(1, 10, by = 1))
)
```

| mtry  |       Accuracy          | 
|-------|-------------------------|
| `1`   | `r rf_mod$result[1,2]`  | 
| `2`   | `r rf_mod$result[2,2]`  |
| `3`   | `r rf_mod$result[3,2]`  |
| `4`   | `r rf_mod$result[4,2]`  |
| `5`   | `r rf_mod$result[5,2]`  |
| `6`   | `r rf_mod$result[6,2]`  |
| `7`   | `r rf_mod$result[7,2]`  |
| `8`   | `r rf_mod$result[8,2]`  |
| `9`   | `r rf_mod$result[9,2]`  |
| `10`  | `r rf_mod$result[10,2]` |

## Exercise 4
```{r}
get_best_result = function(caret_fit) {
  best = which(rownames(caret_fit$results) == rownames(caret_fit$bestTune))
  best_result = caret_fit$results[best, ]
  rownames(best_result) = NULL
  best_result
}
```

```{r}
test_pred1 = predict(knn_mod, newdata = wisc_test)
```

```{r}
test_pred2 = predict(knn_mod2, newdata = wisc_test)
```

```{r}
prob = predict(rf_mod, newdata = wisc_test, type = "prob")[10, 2]
```

```{r}
test_pred3 = predict(rf_mod, newdata = wisc_test)
result = confusionMatrix(test_pred3, wisc_test$class, positive = "M")
```

```{r}
calc_acc = function(actual, predicted) {
  mean(actual == predicted)
}
```

```{r}
calc_acc(actual = wisc_test$class, predicted = test_pred3)
```

| Question |                            Answer                             | 
|----------|-------------------------------------------------------------- |
|  `(a)`   | `r get_best_result(knn_mod)$k`                                |
|  `(b)`   | `r get_best_result(knn_mod)$Accuracy`                         |
|  `(c)`   | `r calc_acc(actual = wisc_test$class, predicted = test_pred1)`|
|  `(d)`   | `r get_best_result(knn_mod2)$k`                               |
|  `(e)`   | `r get_best_result(knn_mod2)$Accuracy`                        |
|  `(f)`   | `r calc_acc(actual = wisc_test$class, predicted = test_pred2)`|
|  `(g)`   | `KNN is performing better with predictor scaling `            |
|  `(h)`   | `r get_best_result(rf_mod)$mtry`                              |
|  `(i)`   | `r prob`                                                      |        
|  `(j)`   | `r result$byClass['Sensitivity']`                             |
|  `(k)`   | `r result$byClass['Specificity'] `                            |
|  `(l)`   | `the random forest is performing better`                      |